\propertitle

### 2019-11-20 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T17:09:08+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? The approach presented here gives probability
distributions over the possible answers to these questions.

@@

The idea behind our approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
  \item\label{item:maxent_step} we build a distribution $\px(\yA)$ for the total
  activity of the full network using the maximum-entropy method;
  \item\label{item:sample_step} the constrained averages used in the
  maximum-entropy method for the full network are, in turn, determined via
  sampling theory from the constrained averages for the sample.
\end{enumerate}
Let's discuss these points in detail.

Regarding~\ref{item:maxent_step}, we assume familiarity with the
maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
brevity. It amounts to a pair of prescriptions: choose the distribution,
among those satisfying specific convex constraints, such as fixed
expectations, that has minimum relative entropy with respect to a reference
distribution, often taken to be the uniform one; and judge those
expectations to be equal to measured averages. We add two remarks about
this method that are seldom made in the literature. First remark: the
distribution $\px(\yA)$ given by this method is the zeroth-order
approximation \citep[in the sense of Laplace's
method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
four different distributions for the full network:
\begin{enumerate}[label=(\roman*)]
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
The maximum-entropy distribution is thus an approximation of our belief
distribution about four different quantities. Note that the four
distributions above numerically differ in higher-order approximations.
Second remark: the maximum-entropy method based on the Shannon entropy
implicitly makes some assumptions about the probabilities for the long-run
frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these two
remarks further in \sect~\ref{sec:assumptions}.

In our case, to apply the maximum-entropy prescriptions to the total
activity of the full network we need to fix some averages of its belief
distribution, for example the distribution's moments. But we don't have any
measured moments for the full network to equate the distribution moments
to. Here enters point~\ref{item:sample_step}: the probability calculus
gives an exact, linear relation between the first $m$ moments for the full
network and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
vice versa at every time bin. This relation is a classical result of
sampling theory
\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
also][]{whitworth1897}.

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-20T11:41:20+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? The approach presented here gives a probability
distribution that approximately answers all these questions.

@@

The idea behind our approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
\item \label{item:sample_step} Using sampling theory we determine the
  relation between some expected values -- specifically, moments -- for the
  total activity $\ya$ of the sample and corresponding expected values $M$
  for the total activity $\yA$ of the full network.
\item \label{item:maxent_step} Using the maximum-entropy method we build a
  distribution $\px(\yA \| M, N)$ for the total activity of the full
  network of $N$ neurons, using as constraints the expected values $M$
  found in the previous step.
\end{enumerate}
We now discuss the ideas behind these steps more in detail, leaving their
precise mathematical implementation to
appendix~\ref{sec:derivation_maxent_fullnet}.

Step~\ref{item:sample_step} is possible because the probability calculus
gives an exact, linear relation between the first $m$ moments for the full
network and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}; the ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA)$ for the full-network activity and its marginal
$\pxx(\ya)$ for the sample activity. It is a classical result of sampling
theory.

***

\citep[\chaps~I--IV]{whitworth1867_r1965}[\chap~II]{feller1950_r1968}[\chap~3]{jaynes1994_r2003}[see
also][]{whitworth1897}.

@@

Regarding~\ref{item:maxent_step}, we assume familiarity with the
maximum-entropy method. We actually use the \emph{minimum-relative-entropy}
method \citep{hobsonetal1973}, but call it \enquote{maximum-entropy} for
brevity. It amounts to a pair of prescriptions: choose the distribution,
among those satisfying specific convex constraints, such as fixed
expectations, that has minimum relative entropy with respect to a reference
distribution, often taken to be the uniform one; and judge those
expectations to be equal to measured averages. We add two remarks about
this method that are seldom made in the literature. First remark: the
distribution $\px(\yA \| M, N)$ given by this method is the zeroth-order
approximation \citep[\mynote{move to sect}in the sense of Laplace's
method:][\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000} of
four different distributions for the full network:
\begin{enumerate}[label=(\roman*)]
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
Note that the four distributions above numerically differ in higher-order
approximations. Second remark: the maximum-entropy method based on the
Shannon entropy implicitly makes some assumptions about the probabilities
for the long-run frequency distributions \citep[\mynote{move to
  sect}]{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
two remarks further in \sect~\ref{sec:assumptions}.

In our case, to apply the maximum-entropy prescription to the total
activity of the full network we need to fix some averages of its belief
distribution, for example the distribution's moments. But we don't have any
measured moments for the full network to equate the distribution moments
to.
